---
title: "Facets & Aspect"
output: html_notebook
---
We can set the aspect ratio of a plot with coord_fixed() or coord_equal(). Both use ratio = 1 as a default. A 1:1 aspect ratio is most appropriate when two continuous variables are on the same scale, as with the iris dataset.

All variables are measured in centimeters, so it only makes sense that one unit on the plot should be the same physical distance on each axis. This gives a more truthful depiction of the relationship between the two variables since the aspect ratio can change the angle of our smoothing line. This would give an erroneous impression of the data.

Complete the basic scatter plot function using the iris data frame to plot Sepal.Width onto the y aesthetic, Sepal.Length onto the x and Species onto col. You should understand all the other functions used in this plot call by now. This is saved in an object called base.plot.

```{r}
base.plot <- ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width, col = Species)) +
               geom_jitter() +
               geom_smooth(method = "lm", se = FALSE)

```

## Polar Coordinates

```{r}
# Create a stacked bar plot: wide.bar
wide.bar <- ggplot(mtcars, aes(x = 1, fill = cyl)) +
              geom_bar()

# Convert wide.bar to pie chart
wide.bar +
  coord_polar(theta = "y")

# Create stacked bar plot: thin.bar
thin.bar <- ggplot(mtcars, aes(x = 1, fill = cyl)) +
              geom_bar(width = 0.1) +
              scale_x_continuous(limits = c(0.5, 1.5))

# Convert thin.bar to "ring" type pie chart
thin.bar + 
  coord_polar(theta = "y")
```

Facets are another way of presenting categorical variables. Recall that we saw all the ways of combining variables, both categorical and continuous, in the aesthetics chapter. Sometimes it's possible to overdo it. Here we'll present a plot with 6 variables and see if we can add even more.

Let's begin by using a trick to map two variables onto two color scales - hue and lightness. We combine cyl and am into a single variable cyl_am. To accommodate this we also make a new color palette with alternating red and blue of increasing darkness. This is saved as myCol. If you are not familiar with these steps, execute the code piece-by-piece.

```{r}
library(RColorBrewer)
mtcars$cyl_am <- paste(mtcars$cyl, mtcars$am, sep = "_")
myCol <- rbind(brewer.pal(9, "Blues")[c(3,6,8)],
               brewer.pal(9, "Reds")[c(3,6,8)])

ggplot(mtcars, aes(x = wt, y = mpg, col = cyl_am)) +
  geom_point() +
  # Add a manual colour scale
  scale_color_manual(values = myCol)
```
When you have a categorical variable with many levels which are not all present in each sub-group of another variable, it may be desirable to drop the unused levels. As an example let's return to the mammalian sleep dataset, mamsleep. It is available in your workspace.

The variables of interest here are name, which contains the full popular name of each animal, and vore, the eating behavior. Each animal can only be classified under one eating habit, so if we facet according to vore, we don't need to repeat the full list in each sub-plot.

```{r}

# Wrangle mamsleep a little
mamsleep <-  msleep %>% 
  select(name, vore, sleep_total, sleep_rem) %>% 
  gather(key = "sleep", val = "time", sleep_total, sleep_rem) %>% 
  na.omit

p <- ggplot(mamsleep, aes(x = time, y = name, col = sleep)) +
  geom_point()
p + facet_grid(vore~., scale = "free_y", space = "free_y")
  
```




