---
title: "ggplot - Datacamp"
output: html_notebook
---

# Smoothing
Welcome to the exercises for the second ggplot2 course!

To practice on the remaining four layers (statistics, coordinates, facets and themes), we'll continue working on several datasets that we already encountered in the first course.

The mtcars dataset contains information for 32 cars from Motor Trends magazine from 1973. This dataset is small, intuitive, and contains a variety of continuous and categorical (both nominal and ordinal) variables.

In the previous course we learned how to effectively use some basic geometries, such as point, bar and line. In the first chapter of this course we'll explore statistics associated with specific geoms, for example, smoothing and lines.

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*. 



Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

Familiarize yourself again with the mtcars dataset using str().

```{r}
str(mtcars)
```

Extend the first ggplot call: add a LOESS smooth to the scatter plot (which is the default) with geom_smooth(). We want to have the actual values and the smooth on the same plot.

```{r}
library(ggplot2)
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() + 
  geom_smooth()
```
Change the previous plot to use an ordinary linear model, by default it will be y ~ x, so you don't have to specify a formula. You should set the method argument to "lm".

```{r}
library(ggplot2)
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() + 
  geom_smooth(method = "lm")
```

Modify the previous plot to remove the 95% CI ribbon. You should set the se argument to FALSE.
```{r}
library(ggplot2)
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() + 
  stat_smooth(method = "lm", se = FALSE)
```

We'll continue with the previous exercise by considering the situation of looking at sub-groups in our dataset. For this we'll encounter the invisible group aesthetic.

 In this ggplot command our smooth is calculated for each subgroup because there is an invisible aesthetic, group which inherits from col.

```{r}
ggplot(mtcars, aes(x = wt, y = mpg, col = factor(cyl))) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE)
```

```{r}
# ggplot2 is already loaded

# 1 - Define cyl as a factor variable
ggplot(mtcars, aes(x = wt, y = mpg, col = factor(cyl))) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE)  


# 2 - Plot 1, plus another stat_smooth() containing a nested aes()

ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point(aes(col=factor(gear))) +
  stat_smooth(method = "lm", se = FALSE, aes(group = 1))


```

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

Plot 1: Recall that LOESS smoothing is a non-parametric form of regression that uses a weighted, sliding-window, average to calculate a line of best fit. We can control the size of this window with the span argument.

```{r}
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  # Add span below
  geom_smooth(se = FALSE, span = 0.7)
```


Plot 1: Recall that LOESS smoothing is a non-parametric form of regression that uses a weighted, sliding-window, average to calculate a line of best fit. We can control the size of this window with the span argument.

Add span, set it to 0.7.

```{r}
ggplot(mtcars, aes(x=wt, y=mpg, col=factor(cyl))) +
  geom_point() + 
  stat_smooth(se = FALSE, span  = 0.7)
```
Plot 2: In this plot, we set a linear model for the entire dataset as well as each subgroup, defined by cyl. In the second stat_smooth(),

Set method to "loess" (this is the default with a small (n < 1000) data set, but we will specify it explicitly).
Add span, set it to 0.7.

```{r}
ggplot(mtcars, aes(x = wt, y = mpg, col = factor(cyl))) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE) +
  stat_smooth(se = FALSE, span = 0.7, aes(group = 1), col = "black")
```
Plot 3: Plot 2 presents a problem because there is a black line on our plot that is not included in the legend. To get this, we need to map something to col as an aesthetic, not just set col as an attribute.

Add col to the aes() function in the second stat_smooth(), set it to "All". This will name the line properly.
Remove the col attribute in the second stat_smooth(). Otherwise, it will overwrite the col aesthetic.


```{r}
ggplot(mtcars, aes(x=wt, y=mpg, col=factor(cyl))) + 
  geom_point() + 
  stat_smooth(se = FALSE) + 
  stat_smooth(aes(group = 1, col = "Todos"), se  = FALSE)
```
Manual scale

```{r}
# Plot 4: Add scale_color_manual to change the colors
library(RColorBrewer)
myColors <- c(brewer.pal(3, "Dark2"), "black")
ggplot(mtcars, aes(x = wt, y = mpg, col = factor(cyl))) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE, span = 0.7) +
  stat_smooth(method = "loess", 
              aes(group = 1, col="All"), 
              se = FALSE, span = 0.7) +
  # Add correct arguments to scale_color_manual
  scale_color_manual("Cylinders", values=myColors)
```

## Modifying stat_smooth (2)

In this exercise we'll take a look at a more subtle example of defining and using linear models. ggplot2 and the Vocab data frame are already loaded for you.

Plot 1: This code produces a jittered plot of vocabulary against education, variables from the Vocab data frame.

```{r}
library(readr)
Vocab <- read_csv("Vocab.csv")
ggplot(Vocab, aes(x = education, y = vocabulary)) + 
  geom_jitter(alpha = 0.2)
```

Add a stat_smooth() layer with method set to "lm". Hide the CI ribbons by using se = FALSE

```{r}
ggplot(Vocab, aes(x=education, y = vocabulary)) +
  geom_jitter() + 
  stat_smooth(method = "lm", se = FALSE)
```

```{r}
# Plot 4: Set a color brewer palette
ggplot(Vocab, aes(x = education, y = vocabulary, col = factor(year))) +
  stat_smooth(method = "lm", se = FALSE) +  # smooth
  scale_color_brewer()  # colors


```

Plot 5: To get the proper colors, we can use col = year, because the variable year is type integer and we want a continuous scale. However, we'll need to specify the invisible group aesthetic so that our linear models are still calculated appropriately. The scale layer, scale_color_gradientn(), has been provided for you - this allows us to map a continuous variable onto a colour scale.

Add group = factor(year) inside aes().
Inside stat_smooth(), set alpha = 0.6 and size = 2.

```{r}
# Plot 5: Add the group aes, specify alpha and size
ggplot(Vocab, aes(x = education, y = vocabulary, col = year, group = factor(year))) +
  stat_smooth(method = "lm", se = FALSE, alpha = 0.6, size = 2) +
  scale_color_gradientn(colors = brewer.pal(9, "YlOrRd"))
```

