---
title: "Stats"
output: html_notebook
---
Calculating statistics
```{r}
set.seed(123)
xx <- rnorm(100)
mean(xx)
sd(xx)
mean(xx) +  sd(xx) * c(-1, 1)

library(Hmisc)
smean.sdl(xx, mult = 1)
# wrapper around smean.sdl
mean_sdl(xx, mult = 1)

## ggplot::stat_summary uses geom_pointrange
## which uses y ymin ymax
ggplot(iris, aes(x = Species, y = Sepal.Length)) +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult=1))

ggplot(iris, aes(x = Species, y = Sepal.Length)) + 
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult=1),
               geom = "errorbar", width = 0.1)
```

Outputs from fun.data must match the arguments expected by the geom we use.
```{r}
foo <- function(x) c(y=4, ymin=1, ymax=6)
ggplot(iris, aes(x = Species, y = Sepal.Length)) +
  stat_summary(fun.data = foo) # uses geom_pointbar by default

ggplot(iris, aes(x = Species, y = Sepal.Length)) +
  stat_summary(fun.data = foo, geom = "point")

ggplot(iris, aes(x = Species, y = Sepal.Length)) +
  stat_summary(fun.data = foo, geom = "errorbar")

bar <- function(x) c(y=4, xmin=1, xmax=5)
# error
# ggplot(iris, aes(x = Species, y = Sepal.Length)) +
#   stat_summary(fun.data = bar, geom = "errorbar")

```


```{r}
library(MASS)
mam.new <- data.frame(body = log10(mammals$body))
ggplot(mam.new, aes(x=body)) + 
  geom_histogram(aes(y = ..density..)) +
  geom_rug() + 
  stat_function(fun = dnorm, colour = "red",
                args = list(mean = mean(mam.new$body),
                          sd = sd(mam.new$body)))
```



In this exercise we'll consider the preparations. That means we'll make sure the data is in the right format and that all the positions that we might use in our plots are defined. Lastly, we'll set the base layer for our plot. ggplot2 is already loaded, so you can get started straight away!

In mtcars, cyl and am are classified as continuous, but they are actually categorical. Previously we just used factor(), but here we'll modify the actual dataset. Change cyl and am to be categorical in the mtcars data frame using as.factor.



```{r}
library(ggplot2)
mtcars2 <- mutate(mtcars, cyl = as.factor(cyl),
                          am = as.factor(am))
# 
```
posn.d, using position_dodge() with a width of 0.1
```{r}
posn.d <- position_dodge(width = 0.1)
posn.jd <- position_jitterdodge(jitter.width = 0.1, dodge.width = 0.2)
posn.j <- position_jitter(width = 0.2)
```

Finally, we'll make our base layers and store it in the object wt.cyl.am. Make the base call for ggplot mapping cyl to the x, wt to y, am to both col and fill. Also set group = am inside aes(). The reason for these redundancies will become clear later on.

```{r}
wt.cyl.am <- ggplot(mtcars2, aes(cyl, wt, col = am, fill = am, group = am))
# Plot 1: Jittered, dodged scatter plot with transparent points
wt.cyl.am +
  geom_point(position = posn.j, alpha = 0.6)

```
```{r}
# wt.cyl.am, posn.d, posn.jd and posn.j are available

# Plot 1: Jittered, dodged scatter plot with transparent points
wt.cyl.am +
  geom_point(position = posn.jd, alpha = 0.6)

# Plot 2: Mean and SD - the easy way
wt.cyl.am + 
stat_summary(fun.data = mean_sdl, fun.args=list(mult=1), position = posn.d)


# Plot 3: Mean and 95% CI - the easy way
wt.cyl.am +
stat_summary(fun.data = mean_cl_normal, position = posn.d)


# Plot 4: Mean and SD - with T-tipped error bars - fill in ___
wt.cyl.am +
  stat_summary(geom = "point", fun.y = mean,
               position = posn.d) +
  stat_summary(geom = "errorbar", fun.data = mean_sdl,
               position = posn.d, fun.args = list(mult = 1), width = 0.1)
```


```{r}
# The base ggplot command; you don't have to change this
wt.cyl.am <- ggplot(mtcars, aes(x = cyl,y = wt, col = am, fill = am, group = am))

# Play vector xx is available

# Function to save range for use in ggplot
gg_range <- function(x) {
  # Change x below to return the instructed values
  data.frame(ymin = min(x), # Min
             ymax = max(x)) # Max
}

gg_range(xx)
# Required output
#   ymin ymax
# 1    1  100

# Function to Custom function
med_IQR <- function(x) {
  # Change x below to return the instructed values
  data.frame(y = median(x), # Median
             ymin = quantile(x)[2], # 1st quartile
             ymax = quantile(x)[4])  # 3rd quartile
}

med_IQR(xx)
# Required output
#        y  ymin  ymax
# 25% 50.5 25.75 75.25

# Add three stat_summary calls to wt.cyl.am
wt.cyl.am +
  stat_summary(geom = "linerange", fun.data = med_IQR,
               position = posn.d, size = 3) +
  stat_summary(geom = "linerange", fun.data = gg_range,
               position = posn.d, size = 3,
               alpha = 0.4) +
  stat_summary(geom = "point", fun.y = median,
               position = posn.d, size = 3,
               col = "black", shape = "X")
```

